Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2023-12-28 paper.tex
Forgot a closing parenthesis
Add a subsection about timings

To detect the direction of anisotropy we use an $N \times N$ (where $N$ is
dimensionality of the sample) covariance matrix for coordinates of points in the
phase of interest:
\begin{equation}
  C_{ij} = \sum_{x \in \left\{
    \begin{array}{l}
      \text{Coordinates of points} \\
      \text{in the phase of interset}
    \end{array}
    \right\}} (x_i - \overline{x_i})(x_j - \overline{x_j})
\end{equation}
and decompose $C$ as $T \Lambda T^{-1}$ where $T$ is a unitary matrix and
$\Lambda$ is diagonal. $T$ can always be made so that it defines a rotation (by
multiplication of one of the columns by $-1$ if needed, so that
$\det T = 1$. Application of this rotation to the sample makes the direction of
anisotropy parallel to one of the axes, so anisotropy can be better featured in
the correlation function \cref{fig:aniso}.
\begin{figure*}[tp]
  \centering
  \subfigure[An original image of an anisotropic pore.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso.png}
    \label{fig:aniso-unrot}}
  \hfill
  \subfigure[A rotated image of the pore, so it is elongated along the axis
    $X$.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso-rot.png}
    \label{fig:aniso-rot}}
  \vskip\baselineskip
  \subfigure[$S_3$ function for the original image]{
    \includegraphics[width=0.3\linewidth]{images/aniso-s3.png}}
  \hfill
  \subfigure[$S_3$ function for the rotated image. A vector $(0.92, -0.4)$ is a
    direction of elongation of the pore.]{
    \includegraphics[width=0.3\linewidth]{images/aniso-rot-s3.png}
    \label{fig:aniso-rot-s3}}
  \hfill
  \subfigure[$S_3$ function averaged across 100 random rotations.]{
    \includegraphics[width=0.3\linewidth]{images/aniso-avg-s3.png}
    \label{fig:aniso-avg}}
  \caption[]{Behaviour of our algorithm for calculation of the $S_3$ function
    when dealing with highly anisotropic media. The pattern \cref{fig:pattern}
    which we use in our algorithm is most sensitive to anisotropy which is
    directed along one of the axes. A result on \cref{fig:aniso-avg} averaged
    across rotations is similar to the algorithm proposed by Malmir~et~al.}
  \label{fig:aniso}
\end{figure*}
The above said holds for every algorithm introduced in the paper.

@@

>>>>

To detect the direction of anisotropy we use an $N \times N$ (where $N$ is
dimensionality of the sample) covariance matrix for coordinates of points in the
phase of interest:
\begin{equation}
  C_{ij} = \sum_{x \in \left\{
    \begin{array}{l}
      \text{Coordinates of points} \\
      \text{in the phase of interset}
    \end{array}
    \right\}} (x_i - \overline{x_i})(x_j - \overline{x_j})
\end{equation}
and decompose $C$ as $T \Lambda T^{-1}$ where $T$ is a unitary matrix and
$\Lambda$ is diagonal. $T$ can always be made so that it defines a rotation (by
multiplication of one of the columns by $-1$ if needed, so that
$\det T = 1$). Application of this rotation to the sample makes the direction of
anisotropy parallel to one of the axes, so anisotropy can be better featured in
the correlation function \cref{fig:aniso}.
\begin{figure*}[tp]
  \centering
  \subfigure[An original image of an anisotropic pore.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso.png}
    \label{fig:aniso-unrot}}
  \hfill
  \subfigure[A rotated image of the pore, so it is elongated along the axis
    $X$.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso-rot.png}
    \label{fig:aniso-rot}}
  \vskip\baselineskip
  \subfigure[$S_3$ function for the original image]{
    \includegraphics[width=0.3\linewidth]{images/aniso-s3.png}}
  \hfill
  \subfigure[$S_3$ function for the rotated image. A vector $(0.92, -0.4)$ is a
    direction of elongation of the pore.]{
    \includegraphics[width=0.3\linewidth]{images/aniso-rot-s3.png}
    \label{fig:aniso-rot-s3}}
  \hfill
  \subfigure[$S_3$ function averaged across 100 random rotations.]{
    \includegraphics[width=0.3\linewidth]{images/aniso-avg-s3.png}
    \label{fig:aniso-avg}}
  \caption[]{Behaviour of our algorithm for calculation of the $S_3$ function
    when dealing with highly anisotropic media. The pattern \cref{fig:pattern}
    which we use in our algorithm is most sensitive to anisotropy which is
    directed along one of the axes. A result on \cref{fig:aniso-avg} averaged
    across rotations is similar to the algorithm proposed by Malmir~et~al.}
  \label{fig:aniso}
\end{figure*}
The above said holds for every algorithm introduced in the paper.

@@

\subsection{Execution times for our algorithms}
In this section we measure execution times of our algorithms for random
3-dimensional arrays of bits (i.e. each element can be either 0 or 1). The
arrays have equal number of elements along each direction therefore
representing cubes of 2-phase porous media. The number of voxels in each cube
increases from $1.25 \cdot 10^5$ to $2.7 \cdot 10^7$ linearly. The time of execution
is measured for each function executed on a CPU and a GPU. Hardware and software
components used for testing are listed in \cref{tab:machine}.
\begin{table*}[!htp]
  \centering
  \begin{tabular}{|c|c|}
    CPU & Intel Celeron 333 MHz \\
    RAM & 100500 GB DDR4 PC-XXXX \\
    GPU & AMD Navi 21 \\
    Julia & 1.10.0 \\
    CUDA.jl & 4.0
  \end{tabular}
  \caption{Parameters of the machine used to measure execution times of the
    algorithms described in this paper.}
  \label{tab:machine}
\end{table*}
\begin{figure*}[tp]
  \centering
  \subfigure[CPU]{
    \includegraphics[width=0.4\linewidth]{images/timings-cpu.png}}
  \hfill
  \subfigure[GPU]{
    \includegraphics[width=0.4\linewidth]{images/timings-gpu.png}}
  \caption[]{Execution times of the algorithms proposed in this paper.}
  \label{fig:timings}
\end{figure*}
The results are on \cref{fig:timings}. It can be observed that the time of
execution is a power function of number of voxels. GPU gives a
10x-100x performance boost depending on the correlation function.
